Does language control differ between younger and older adults when switching between versus within languages?

### 2023-05-30 manuscriptINSPIRE.tex
Initial upload of manuscript and data

---
title: Does language control differ between younger and older adults when switching
  between versus within languages?
author: "Angela de Bruin*$^1$, Heidi Kressel$^2$, & Daisy Hemmings$^2$"
output:
   #html_document: default
   bookdown::pdf_document2:
    keep_tex: true
    toc: false
    fig_caption: yes
   pdf_document: default

header-includes:
  \newcommand{\beginsupplement}{\setcounter{table}{0}       
  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} 
  \renewcommand{\thefigure}{S\arabic{figure}}}
---

``` {r setup, include=FALSE}

# flag to specify if analyses are done from scratch (takes ~6 minutes)
# if set to 0, precomputed analyses are loaded from an RData file
doanalysis <- 1

#Note: data were analysed using lme4 package version 1.1.33 and lmerTest 3.1.3 in R 4.3.0

packagelist <- c('lme4','lmerTest','tidyverse','ggplot2','rmarkdown','bookdown','kableExtra','rstatix')
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

#devtools::install_github("kupietz/kableExtra")

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

``` {r trimmingfunction, include=FALSE}

#------------------------------------------------------------------------------
#' RT trimming with standard deviation criterion (Grange, 2017, sdtrim package)
#'
#' \code{sdTrim} takes a data frame of RT data and returns trimmed rt
#' data that fall below a set set criterion (based on standard deviations
#' above a particular mean). The criterion can be based on the mean of the
#' whole set of data, based on the mean per experimental condition, based on
#' the mean per participant, or based on the mean of each participant in each
#' experimental condition.
#'
#' By passing a data frame containing raw response time data, together with
#' trimming criteria, the function will return trimmed data, either in the form
#' of trial-level data or in the form of means/medians for each subject &
#' condition.
#'
#' @param data A data frame. It must contain columns named "participant",
#' "condition", "rt", and "accuracy". The RT can be in seconds
#' (e.g., 0.654) or milliseconds (e.g., 654). Condition will consist
#' of strings. "accuracy" must be 1 for correct and 0 for error
#' responses.
#' @param minRT The lower criteria for acceptable response time. Must be in
#' the same form as rt column in data frame (e.g., in seconds OR milliseconds).
#' All RTs below this value are removed before proceeding with SD trimming.
#' @param sd The upper criteria for standard deviation cut-off.
#' @param perCondition Set to TRUE if the user wishes the trimming to occur per
#' condition of the experimental design.
#' @param perParticipant Set to TRUE if the user wishes the trimming to occur
#' per participant.
#' @param omitErrors If set to TRUE, error trials will be removed before
#' conducting trimming procedure. Final data returned will not be influenced
#' by errors in this case.
#' @param returnType Request nature of returned data. "raw" returns trial-
#' level data excluding trimmed data; "mean" returns mean response times per
#' participant for each experimental condition identified; "median" returns
#' median response times per participant for each experimental condition
#' identified.
#' @param digits How many decimal places to round to after trimming?
#' @examples
#' # load the example data that ships with trimr
#' data(exampleData)
#'
#' # perform the trimming with SD trimming per condition, returning mean RT
#' trimmedData <- sdTrim(data = exampleData, minRT = 150, sd = 2.5,
#' perCondition = TRUE, perParticipant = FALSE, returnType = "mean")
#'
#' @importFrom stats median sd
#'
#' @export
sdTrim <- function(data, minRT, sd, perCondition = TRUE, perParticipant = TRUE,
                   omitErrors = TRUE, returnType = "mean", digits = 3){

  ###-------------
  if(perCondition == FALSE & perParticipant == FALSE){
    # change the variable name for sd (as this is an R function)
    stDev <- sd

    # remove errors if the user has asked for it
    if(omitErrors == TRUE){
      trimmedData <- subset(data, data$accuracy == 1)
    } else {
      trimmedData <- data
    }

    # get the list of participant numbers
    participant <- sort(unique(trimmedData$participant))

    # get the list of experimental conditions
    conditionList <- unique(trimmedData$condition)

    # trim the data to remove trials below minRT
    trimmedData <- subset(trimmedData, trimmedData$rt > minRT)

    # what is the mean & SD of the whole group's data?
    meanRT <- mean(trimmedData$rt)
    sdRT <- sd(trimmedData$rt)

    # what is the cut-off value?
    cutoff <- meanRT + (stDev * sdRT)

    # remove these rts
    trimmedData <- subset(trimmedData, trimmedData$rt < cutoff)

    # if the user asked for trial-level data, return immediately to user
    if(returnType == "raw"){
      return(trimmedData)
    }

    # if the user has asked for means, then split the data into separate
    # conditions, and display the means per condition.
    if(returnType == "mean"){

      # ready the final data set
      finalData <- matrix(0, nrow = length(participant),
                          ncol = length(conditionList))

      # give the columns the condition names
      colnames(finalData) <- conditionList

      # add the participant column
      finalData <- cbind(participant, finalData)

      # convert to data frame
      finalData <- data.frame(finalData)

      # loop over all conditions, and over all subjects, and find mean RT

      j <- 2 # to keep track of conditions looped over. Starts at 2 as this is
      # where the first condition's column is.

      for(currCondition in conditionList){

        # get the current condition's data
        tempData <- subset(trimmedData, trimmedData$condition == currCondition)

        #now loop over all participants
        i <- 1

        for(currParticipant in participant){

          # get that participant's data
          participantData <- subset(tempData,
                                    tempData$participant == currParticipant)

          # calculate & store their mean response time
          finalData[i, j] <- round(mean(participantData$rt), digits = digits)

          # update participant counter
          i <- i + 1
        }

        # update nCondition counter
        j <- j + 1

      } # end of condition loop

      return(finalData)

    } ## end MEAN sub-function

    # if the user has asked for medians, then split the data into separate
    # conditions, and display the medians per condition.
    if(returnType == "median"){

      # ready the final data set
      finalData <- matrix(0, nrow = length(participant),
                          ncol = length(conditionList))

      # give the columns the condition names
      colnames(finalData) <- conditionList

      # add the participant column
      finalData <- cbind(participant, finalData)

      # convert to data frame
      finalData <- data.frame(finalData)

      # loop over all conditions, and over all subjects, and find mean RT

      j <- 2 # to keep track of conditions looped over. Starts at 2 as this is
      # where the first condition's column is.

      for(currCondition in conditionList){

        # get the current condition's data
        tempData <- subset(trimmedData, trimmedData$condition == currCondition)

        #now loop over all participants
        i <- 1

        for(currParticipant in participant){

          # get that participant's data
          participantData <- subset(tempData,
                                    tempData$participant == currParticipant)

          # calculate & store their mean response time
          finalData[i, j] <- round(median(participantData$rt), digits = digits)

          # update participant counter
          i <- i + 1
        }

        # update nCondition counter
        j <- j + 1

      } # end of condition loop

      return(finalData)
    }

  } # end of perCell == FALSE & perParticipant == FALSE

  ###-------------
  if(perCondition == TRUE & perParticipant == FALSE){

    # change the variable name for sd (as this is an R function)
    stDev <- sd

    # remove errors if the user has asked for it
    if(omitErrors == TRUE){
      trimmedData <- subset(data, data$accuracy == 1)
    } else {
      trimmedData <- data
    }

    # get the list of participant numbers
    participant <- sort(unique(trimmedData$participant))

    # get the list of experimental conditions
    conditionList <- unique(trimmedData$condition)

    # trim the data to remove trials below minRT
    trimmedData <- subset(trimmedData, trimmedData$rt > minRT)

    ### do "raw"
    if(returnType == "raw"){

      # initialise variable to keep trimmed data in
      finalData <- NULL

      # loop over each condition
      for(cond in conditionList){

        # get the data, & find cutoff
        curData <- subset(trimmedData, trimmedData$condition == cond)
        curMean <- mean(curData$rt)
        curSD <- sd(curData$rt)
        curCutoff <- curMean + (stDev * curSD)
        curCutoff2 <- curMean - (stDev * curSD)

        # trim the data
        curData <- subset(curData, curData$rt < curCutoff & curData$rt > curCutoff2)

        # bind the data
        finalData <- rbind(finalData, curData)
      }

      return(finalData)
    }

    ### do "mean"
    if(returnType == "mean"){

      ## first, find the cutoff for each condition, and remove the necessary
      ## trials

      # initialise variable to keep trimmed data in
      tempData <- NULL

      for(cond in conditionList){
        # get the data, & find cutoff
        curData <- subset(trimmedData, trimmedData$condition == cond)
        curMean <- mean(curData$rt)
        curSD <- sd(curData$rt)
        curCutoff <- curMean + (stDev * curSD)

        # trim the data
        curData <- subset(curData, curData$rt < curCutoff)

        # bind the data
        tempData <- rbind(tempData, curData)
      }

      # change variable names
      trimmedData <- tempData
      tempData <- NULL

      ## now loop over each subject and calculate their average
      # ready the final data set
      finalData <- matrix(0, nrow = length(participant),
                          ncol = length(conditionList))

      # give the columns the condition names
      colnames(finalData) <- conditionList

      # add the participant column
      finalData <- cbind(participant, finalData)

      # convert to data frame
      finalData <- data.frame(finalData)

      # loop over conditions & subjects and calculate their average

      # to index over conditions. It starts at 2 because this is the first
      # column in the data frame containing condition information
      j <- 2

      for(curCondition in conditionList){

        # get the current condition's data
        tempData <- subset(trimmedData, trimmedData$condition == curCondition)

        #now loop over all participants
        i <- 1

        for(currParticipant in participant){

          # get that participant's data
          participantData <- subset(tempData,
                                    tempData$participant == currParticipant)

          # calculate & store their mean response time
          finalData[i, j] <- round(mean(participantData$rt), digits = digits)

          # update participant counter
          i <- i + 1
        }

        # update nCondition counter
        j <- j + 1
      }

      return(finalData)
    }

  } # end of perCell == TRUE & perParticipant == FALSE

  ###-------------
  if(perCondition == FALSE & perParticipant == TRUE){

    # change the variable name for sd (as this is an R function)
    stDev <- sd

    # remove errors if the user has asked for it
    if(omitErrors == TRUE){
      trimmedData <- subset(data, data$accuracy == 1)
    } else {
      trimmedData <- data
    }

    # get the list of participant numbers
    participant <- sort(unique(trimmedData$participant))

    # get the list of experimental conditions
    conditionList <- unique(trimmedData$condition)

    # trim the data to remove trials below minRT
    trimmedData <- subset(trimmedData, trimmedData$rt > minRT)

    ### do "raw"
    if(returnType == "raw"){

      # initialise variable to keep trimmed data in
      finalData <- NULL

      # loop over each subject
      for(currSub in participant){

        # get the current subject's data
        curData <- subset(trimmedData, trimmedData$participant == currSub)

        # find their mean, sd, & cutoff
        curMean <- mean(curData$rt)
        curSD <- sd(curData$rt)
        curCutoff <- curMean + (stDev * curSD)

        # trim the data
        curData <- subset(curData, curData$rt < curCutoff)

        # bind the data
        finalData <- rbind(finalData, curData)
      }

      return(finalData)
    }

    ### do "mean"
    if(returnType == "mean"){

      # initialise variable to keep trimmed data in
      tempData <- NULL

      # loop over each subject
      for(currSub in participant){

        # get the current subject's data
        curData <- subset(trimmedData, trimmedData$participant == currSub)

        # find their mean, sd, & cutoff
        curMean <- mean(curData$rt)
        curSD <- sd(curData$rt)
        curCutoff <- curMean + (stDev * curSD)

        # trim the data
        curData <- subset(curData, curData$rt < curCutoff)

        # bind the data
        tempData <- rbind(tempData, curData)
      }

      # change variable names
      trimmedData <- tempData
      tempData <- NULL

      # ready the final data set
      finalData <- matrix(0, nrow = length(participant),
                          ncol = length(conditionList))

      # give the columns the condition names
      colnames(finalData) <- conditionList

      # add the participant column
      finalData <- cbind(participant, finalData)

      # convert to data frame
      finalData <- data.frame(finalData)

      # loop over conditions & subjects and calculate their average

      # to index over conditions. It starts at 2 because this is the first
      # column in the data frame containing condition information
      j <- 2

      for(curCondition in conditionList){

        # get the current condition's data
        tempData <- subset(trimmedData, trimmedData$condition == curCondition)

        #now loop over all participants
        i <- 1

        for(currParticipant in participant){

          # get that participant's data
          participantData <- subset(tempData,
                                    tempData$participant == currParticipant)

          # calculate & store their mean response time
          finalData[i, j] <- round(mean(participantData$rt), digits = digits)

          # update participant counter
          i <- i + 1
        }

        # update nCondition counter
        j <- j + 1
      }

      return(finalData)

    }

    ### do "median"
    if(returnType == "median"){

      # initialise variable to keep trimmed data in
      tempData <- NULL

      # loop over each subject
      for(currSub in participant){

        # get the current subject's data
        curData <- subset(trimmedData, trimmedData$participant == currSub)

        # find their mean, sd, & cutoff
        curMean <- mean(curData$rt)
        curSD <- sd(curData$rt)
        curCutoff <- curMean + (stDev * curSD)

        # trim the data
        curData <- subset(curData, curData$rt < curCutoff)

        # bind the data
        tempData <- rbind(tempData, curData)
      }

      # change variable names
      trimmedData <- tempData
      tempData <- NULL

      # ready the final data set
      finalData <- matrix(0, nrow = length(participant),
                          ncol = length(conditionList))

      # give the columns the condition names
      colnames(finalData) <- conditionList

      # add the participant column
      finalData <- cbind(participant, finalData)

      # convert to data frame
      finalData <- data.frame(finalData)

      # loop over conditions & subjects and calculate their average

      # to index over conditions. It starts at 2 because this is the first
      # column in the data frame containing condition information
      j <- 2

      for(curCondition in conditionList){

        # get the current condition's data
        tempData <- subset(trimmedData, trimmedData$condition == curCondition)

        #now loop over all participants
        i <- 1

        for(currParticipant in participant){

          # get that participant's data
          participantData <- subset(tempData,
                                    tempData$participant == currParticipant)

          # calculate & store their median response time
          finalData[i, j] <- round(median(participantData$rt), digits = digits)

          # update participant counter
          i <- i + 1
        }

        # update nCondition counter
        j <- j + 1
      }

      return(finalData)
    }

  } # end of perCell == FALSE & perParticipant == TRUE

  ###-------------
  if(perCondition == TRUE & perParticipant == TRUE){
    # change the variable name for sd (as this is an R function)
    stDev <- sd

    # remove errors if the user has asked for it
    if(omitErrors == TRUE){
      trimmedData <- subset(data, data$accuracy == 1)
    } else {
      trimmedData <- data
    }

    # get the list of participant numbers
    participant <- sort(unique(trimmedData$participant))

    # get the list of experimental conditions
    conditionList <- unique(trimmedData$condition)

    # trim the data to remove trials below minRT
    trimmedData <- subset(trimmedData, trimmedData$rt > minRT)

    ### do "raw"
    if(returnType == "raw"){

      # initialise variable to keep trimmed data in
      finalData <- NULL

      # loop over all participants
      for(currSub in participant){

        # loop over all conditions
        for(currCond in conditionList){

          # get the relevant data
          tempData <- subset(trimmedData, trimmedData$condition == currCond &
                               trimmedData$participant == currSub)

          # find the cutoff
          curMean <- mean(tempData$rt)
          curSD <- sd(tempData$rt)
          curCutoff <- curMean + (stDev * curSD)

          # perform the trim
          curData <- subset(tempData, tempData$rt < curCutoff)

          # store the data
          finalData <- rbind(finalData, curData)
        }
      }

      return(finalData)
    }

    ### do "mean"
    if(returnType == "mean"){

      # ready the final data set
      finalData <- matrix(0, nrow = length(participant),
                          ncol = length(conditionList))

      # give the columns the condition names
      colnames(finalData) <- conditionList

      # add the participant column
      finalData <- cbind(participant, finalData)

      # convert to data frame
      finalData <- data.frame(finalData)

      # intialise looping variable for subjects
      i <- 1

      # loop over all subjects
      for(currSub in participant){

        # intialise looping variable for conditions. It starts at 2 because the
        # first column in the data file containing condition information is the
        # second one.
        j <- 2

        # loop over all conditions
        for(currCond in conditionList){

          # get the relevant data
          tempData <- subset(trimmedData, trimmedData$participant == currSub &
                               trimmedData$condition == currCond)

          # find the cutoff
          curMean <- mean(tempData$rt)
          curSD <- sd(tempData$rt)
          curCutoff <- curMean + (stDev * curSD)

          # trim the data
          curData <- subset(tempData, tempData$rt < curCutoff)

          # find the average, and add to the data frame
          finalData[i, j] <- round(mean(curData$rt), digits = digits)

          # update condition loop counter
          j <- j + 1
        }

        # update participant loop counter
        i <- i + 1
      }

      return(finalData)

    }

    ### do "median"
    if(returnType == "median"){

      # ready the final data set
      finalData <- matrix(0, nrow = length(participant),
                          ncol = length(conditionList))

      # give the columns the condition names
      colnames(finalData) <- conditionList

      # add the participant column
      finalData <- cbind(participant, finalData)

      # convert to data frame
      finalData <- data.frame(finalData)

      # intialise looping variable for subjects
      i <- 1

      # loop over all subjects
      for(currSub in participant){

        # intialise looping variable for conditions. It starts at 2 because the
        # first column in the data file containing condition information is the
        # second one.
        j <- 2

        # loop over all conditions
        for(currCond in conditionList){

          # get the relevant data
          tempData <- subset(trimmedData, trimmedData$participant == currSub &
                               trimmedData$condition == currCond)

          # find the cutoff
          curMean <- mean(tempData$rt)
          curSD <- sd(tempData$rt)
          curCutoff <- curMean + (stDev * curSD)

          # trim the data
          curData <- subset(tempData, tempData$rt < curCutoff)

          # find the average, and add to the data frame
          finalData[i, j] <- round(median(curData$rt), digits = digits)

          # update condition loop counter
          j <- j + 1
        }

        # update participant loop counter
        i <- i + 1
      }

      return(finalData)
    }

  } # end of perCell == TRUE & perParticipant == TRUE

} # end of function

#------------------------------------------------------------------------------

```

``` {r expt1analysis, include=FALSE}

if (doanalysis==1){
#################################
# Exp 1: linguistic noun-verb   #
################################
#open file for linguistic switching task
LS_Exp1 <- read.csv("data/LinguisticSwitching_All.csv")

#### run accuracy analyses first - linguistic noun-verb Exp1 ####
#we are working with DV accuracy01, which just scores accuracy as 0 or 1. Detailed accuracy types are given in "accuracy": 0 = no response; 2 = wrong word; 3 = wrong language; 5 = combination of rules. 1 (correct) and 7 (correct response preceded by incorrect rule trial or break) both count as correct responses.

#### Prepare file for accuracy analysis #### 
#remove first trial after a break in dual-rule part, as trial type cannot be determined 
LS_Exp1 <-  subset(LS_Exp1, LS_Exp1$TrialType != "X")
#contrast code RULE
LS_Exp1$Rule_coded <- as.factor(LS_Exp1$Rule)
LS_Exp1$Rule_coded <- ifelse(LS_Exp1$Rule_coded == "OBJECT", -0.5, ifelse(LS_Exp1$Rule_coded== "ACTION", 0.5, 99))
aggregate(Rule_coded~Rule, data=LS_Exp1, mean, na.rm=TRUE)

#contrast code AGE GROUP
LS_Exp1$Agegroup_coded <- as.factor(LS_Exp1$AgeGroup)
LS_Exp1$Agegroup_coded <- ifelse(LS_Exp1$Agegroup_coded == "Younger", -0.5, ifelse(LS_Exp1$Agegroup_coded== "Older", 0.5, 99))
aggregate(Agegroup_coded~AgeGroup, data=LS_Exp1, mean, na.rm=TRUE)

#### take subset of switch and non-switch trials for switching analysis #### 
LS_Exp1_switching <- subset(LS_Exp1, LS_Exp1$TrialType == "switch" | LS_Exp1$TrialType == "nonswitch")
#contrast code SWITCHING
LS_Exp1_switching$Switching_coded <- as.factor(LS_Exp1_switching$TrialType)
LS_Exp1_switching$Switching_coded <- ifelse(LS_Exp1_switching$Switching_coded == "nonswitch", -0.5, ifelse(LS_Exp1_switching$Switching_coded== "switch", 0.5, 99))
aggregate(Switching_coded~TrialType, data=LS_Exp1_switching, mean, na.rm=TRUE)

#### accuracy analysis - switching ####
#GLME: DV accuracy01, predictors age group, rule, and trial type (switching)

# COMMENTED OUT MODELS THAT DO NOT CONVERGE AND ARE NOT USED IN SUBSEQUENT ANALYSIS
# glme_LS_accSwitching_Exp1<- glmer(accuracy01~ Switching_coded*Agegroup_coded*Rule_coded + (1+Switching_coded*Rule_coded|participant) + (1+Switching_coded*Agegroup_coded*Rule_coded|Image), data = LS_Exp1_switching, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
# #does not converge, remove correlations
# glme_LS_accSwitching_Exp1_1<- glmer(accuracy01~ Switching_coded*Agegroup_coded*Rule_coded + (1+Switching_coded*Rule_coded||participant) + (1+Switching_coded*Agegroup_coded*Rule_coded||Image), data = LS_Exp1_switching, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
# #does not converge, remove item slopes for switching, switching x age group, and switching x age group x rule
# glme_LS_accSwitching_Exp1_2<- glmer(accuracy01~ Switching_coded*Agegroup_coded*Rule_coded + (1+Switching_coded*Rule_coded||participant) + (1+Agegroup_coded+Rule_coded+Switching_coded:Rule_coded+Agegroup_coded:Rule_coded||Image), data = LS_Exp1_switching, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
# #does not converge, remove item slope for rule x switching
# glme_LS_accSwitching_Exp1_3<- glmer(accuracy01~ Switching_coded*Agegroup_coded*Rule_coded + (1+Switching_coded*Rule_coded||participant) + (1+Agegroup_coded+Rule_coded+Agegroup_coded:Rule_coded||Image), data = LS_Exp1_switching, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
# #does not converge, remove item slope for  age group
# glme_LS_accSwitching_Exp1_4<- glmer(accuracy01~ Switching_coded*Agegroup_coded*Rule_coded + (1+Switching_coded*Rule_coded||participant) + (1+Rule_coded+Agegroup_coded:Rule_coded||Image), data = LS_Exp1_switching, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
# #does not converge, remove item slope for  rule
# glme_LS_accSwitching_Exp1_5<- glmer(accuracy01~ Switching_coded*Agegroup_coded*Rule_coded + (1+Switching_coded*Rule_coded||participant) + (1+Agegroup_coded:Rule_coded||Image), data = LS_Exp1_switching, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
# #does not converge, remove item slope for age x rule
# glme_LS_accSwitching_Exp1_6<- glmer(accuracy01~ Switching_coded*Agegroup_coded*Rule_coded + (1+Switching_coded*Rule_coded||participant) + (1|Image), data = LS_Exp1_switching, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
#does not converge, remove participant slope for switching x rule

glme_LS_accSwitching_Exp1_7<- glmer(accuracy01~ Switching_coded*Agegroup_coded*Rule_coded + (1+Switching_coded+Rule_coded||participant) + (1|Image), data = LS_Exp1_switching, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
summary(glme_LS_accSwitching_Exp1_7)

#### mixing effect, accuracy ####
#take subset of single-task and non-switch trials for mixing analysis
LS_Exp1_mixing <- subset(LS_Exp1, LS_Exp1$TrialType == "SingleNaming" | LS_Exp1$TrialType == "nonswitch")
#contrast code MIXING
LS_Exp1_mixing$Mixing_coded <- as.factor(LS_Exp1_mixing$TrialType)
LS_Exp1_mixing$Mixing_coded <- ifelse(LS_Exp1_mixing$Mixing_coded == "SingleNaming", -0.5, ifelse(LS_Exp1_mixing$Mixing_coded== "nonswitch", 0.5, 99))
aggregate(Mixing_coded~TrialType, data=LS_Exp1_mixing, mean, na.rm=TRUE)

#### accuracy analysis - mixing ####
#GLME: DV accuracy01, predictors age group, rule, and trial type (mixing)

# COMMENTED OUT MODELS THAT DO NOT CONVERGE AND ARE NOT USED IN SUBSEQUENT ANALYSIS
# glme_LS_accMixing_Exp1<- glmer(accuracy01~ Mixing_coded*Agegroup_coded*Rule_coded + (1+Mixing_coded*Rule_coded|participant) + (1+Mixing_coded*Agegroup_coded*Rule_coded|Image), data = LS_Exp1_mixing, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
# #does not converge, remove correlations
#glme_LS_accMixing_Exp1_1<- glmer(accuracy01~ Mixing_coded*Agegroup_coded*Rule_coded + (1+Mixing_coded*Rule_coded||participant) + (1+Mixing_coded*Agegroup_coded*Rule_coded||Image), data = LS_Exp1_mixing, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
#does not converge, remove item slopes for mixing x age 
#glme_LS_accMixing_Exp1_2<- glmer(accuracy01~ Mixing_coded*Agegroup_coded*Rule_coded + (1+Mixing_coded*Rule_coded||participant) + (1+Agegroup_coded+Rule_coded+Mixing_coded+Mixing_coded:Rule_coded+Agegroup_coded:Rule_coded+Mixing_coded:Agegroup_coded:Rule_coded||Image), data = LS_Exp1_mixing, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
#does not converge, remove item slope for mixing
#glme_LS_accMixing_Exp1_3<- glmer(accuracy01~ Mixing_coded*Agegroup_coded*Rule_coded + (1+Mixing_coded*Rule_coded||participant) + (1+Agegroup_coded+Rule_coded+Mixing_coded:Rule_coded+Agegroup_coded:Rule_coded+Mixing_coded:Agegroup_coded:Rule_coded||Image), data = LS_Exp1_mixing, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
#does not converge, remove item slopes for mixing x age x rule
glme_LS_accMixing_Exp1_4<- glmer(accuracy01~ Mixing_coded*Agegroup_coded*Rule_coded + (1+Mixing_coded*Rule_coded||participant) + (1+Agegroup_coded+Rule_coded+Mixing_coded:Rule_coded+Agegroup_coded:Rule_coded||Image), data = LS_Exp1_mixing, family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
summary(glme_LS_accMixing_Exp1_4)

#### reaction time Exp 1 linguistic noun-verb #### 
#### outlier removal steps ####

#keep accurate responses only for RT analysis + exclude trials preceded by break or mistake
LS_Exp1_acc <- subset(LS_Exp1, LS_Exp1$accuracy=="1")

#check distribution raw RT
# hist(LS_Exp1_acc$rt)

#log RT for outlier removal
LS_Exp1_acc$rawrt <-LS_Exp1_acc$rt
LS_Exp1_acc$rt <- log(LS_Exp1_acc$rt)
#check histogram log RT
# hist(LS_Exp1_acc$rt)

#create new condition variable that combines participant, trial type, and rule info
LS_Exp1_acc$condition<-paste0(as.character(LS_Exp1_acc$participant), (LS_Exp1_acc$TrialType), (LS_Exp1_acc$Rule))

#remove RT outliers 2.5 SD below/above mean per participant and condition
LS_Exp1_acc_outliers <- sdTrim(data = LS_Exp1_acc, minRT = 0, sd = 2.5, 
                       perCondition = TRUE, perParticipant = FALSE, 
                       returnType = "raw", digits = 20)

#Check distribution of log RT after outlier removal
# hist(LS_Exp1_acc_outliers$rt)

#Write the trimmed data to a .csv file
# write.csv(LS_Exp1_acc_outliers, "LS_Exp1_acc_outliers.csv")

#### Prepare file for RT analysis ####
#### take subset of switch and non-switch trials for switching analysis #### 
LS_Exp1_acc_outliers_switching <- subset(LS_Exp1_acc_outliers, LS_Exp1_acc_outliers$TrialType == "switch" | LS_Exp1_acc_outliers$TrialType == "nonswitch")
#contrast code SWITCHING
LS_Exp1_acc_outliers_switching$Switching_coded <- as.factor(LS_Exp1_acc_outliers_switching$TrialType)
LS_Exp1_acc_outliers_switching$Switching_coded <- ifelse(LS_Exp1_acc_outliers_switching$Switching_coded == "nonswitch", -0.5, ifelse(LS_Exp1_acc_outliers_switching$Switching_coded== "switch", 0.5, 99))
aggregate(Switching_coded~TrialType, data=LS_Exp1_acc_outliers_switching, mean, na.rm=TRUE)

#### RT analysis - switching ####

# COMMENTED OUT MODELS THAT DO NOT CONVERGE AND ARE NOT USED IN SUBSEQUENT ANALYSIS
# #LME: log RT DV, predictors age group, rule, and trial type (switching)
# lme_LS_Exp1<- lmer(rt~ Switching_coded*Agegroup_coded*Rule_coded + (1+Switching_coded*Rule_coded|participant) + (1+Switching_coded*Agegroup_coded*Rule_coded|Image), REML = TRUE, data = LS_Exp1_acc_outliers_switching,control = lmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
# #does not converge, remove correlations
# lme_LS_Exp1_1<- lmer(rt~ Switching_coded*Agegroup_coded*Rule_coded + (1+Switching_coded*Rule_coded||participant) + (1+Switching_coded*Agegroup_coded*Rule_coded||Image), REML = TRUE, data = LS_Exp1_acc_outliers_switching,control = lmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1000000)))
# #does not converge. We first remove item slopes explaining the lowest variance, namely switching x age group

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Does language control differ between younger and older adults when switching between versus within languages?},
  pdfauthor={Angela de Bruin*\^{}1, Heidi Kressel\^{}2, \& Daisy Hemmings\^{}2},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Does language control differ between younger and older adults when switching between versus within languages?}
\author{Angela de Bruin*\(^1\), Heidi Kressel\(^2\), \& Daisy Hemmings\(^2\)}
\date{}

\begin{document}
\maketitle

\(^1\)Department of Psychology, University of York, York, United Kingdom\\
\(^2\)Hull York Medical School, University of York, York, United Kingdom

\textbf{Corresponding author:}\\
Angela de Bruin\\
Department of Psychology\\
University of York\\
York YO10 5DD\\
United Kingdom
